---
title: "Ecosystem Information Review and Strategy Development for More Resilient Fisheries"
subtitle: "SAFMC Project Oversight Team <br />11 September 2025<br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br />"
author: "Sarah Gaichas, Hydra Scientific LLC"
output:
  xaringan::moon_reader:
    css: ["default", "libs/HS_theme3_508_16by9.css"]
    lib_dir: libs
    nature:
      titleSlideClass: ["right", "bottom", "my-title"]
      ratio: "16:9"
      highlightStyle: githubp
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "libs/macros.js"
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_plugins: ["notes", "search"]
---
class: top, left

```{r setup, include=FALSE}
#<span style="font-weight:normal;"><font size="-0.5">$^1$NOAA NMFS Northeast Fisheries Science Center, Woods Hole, MA, USA; $^2$Ocean Associates Inc, Arlington, VA, USA;</font></span>

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = F,
                      fig.retina = 3,
                      warning = F,
                      message = F,
                      error = F)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
#Plotting and data libraries
library(tidyverse)
theme_set(theme_bw())
library(tidyr)
library(here)
library(sf)
library(patchwork)
library(ggpubr)
library(data.table)
library(forcats)
```

```{r, load_refs, include=FALSE, cache=FALSE}
#trick to downlaod bib file from other repo to keep just 1 up to date

# url <- "https://raw.githubusercontent.com/sgaichas/bluefishdiet/main/FishDiet_EcoIndicators.bib"
#   download.file(url, destfile = "./FishDiet_Ecoindicators.bib")

library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "authoryear",
           longnamesfirst = FALSE,
           max.names = 1,
           style = "markdown")
myBib <- ReadBib("./operationalSOE.bib", check = FALSE)

#,
#           hyperlink = FALSE,
#           dashed = FALSE

```

# Project Overview

1) Review the different types of ecosystem data and information including data products and
decision-support tools (e.g. Ecosystem Status Reports, Climate Vulnerability Assessments, risk
assessments) currently available to the regional management councils and evaluate how they are
used in council management processes;

2) Identify opportunities and methods for incorporating identified ecosystem information into
SAFMC management processes, including inter-jurisdictional management decision-making
processes, and identify practical requirements for successful implementation (e.g. data quality,
frequency of information updates, regional council process consistency, implementation
timelines); and

3) Identify opportunities to continue to expand cooperative, constituent-engaged data collection
and research to improve the available ecosystem information in the South Atlantic region (e.g.,
study fleets, cooperative research, citizen science).

---
# Workplan: Objective 1

*Review the different types of ecosystem data and information including data products and decision-support tools (e.g. Ecosystem Status Reports, Climate Vulnerability Assessments, risk assessments) currently available to the regional management councils and evaluate how they are used in council management processes*

.pull-left[
Literature search: *underway*

* Council documents (policy guidance documents, FMPs/FEPs)
* Data products (Ecosystem Status Reports, Climate Vulnerability Assessments)
* Peer-reviewed literature

]

.pull-right[

Structured interviews: *planning*

* Focus on key Council staff and members (2-4 per region)
* Interview Template:
    * Review/validate Data products and uses of record from lit review
    * What other uses are not recorded?
    * Successes and challenges with current information and processes?
    * What might be done differently to better meet Council objectives?

]

---
# Workplan: Objective 2

*Identify opportunities and methods for incorporating identified ecosystem information into SAFMC management processes, including inter-jurisdictional management decision- making processes, and identify practical requirements for successful implementation (e.g. data quality, frequency of information updates, regional council process consistency, implementation timelines)*

.pull-left[
Initial focus on catch specification

* Utility of an ecosystem profile approach?
* Climate vulnerability consideration in uncertainty buffers for ABC setting?
* Risk table approaches used/piloted in other regions?

]

.pull-right[
Broaden to other processes using a review of 3-5 years of Council actions 

* Classify management decisions by data needs and process timelines
* Develop a prioritized list of processes linked to ecosystem information
* Recommend practical pathways for implementation

]

---
# Examples: Objective 2

Case study: risk assessment

Case study: prototype ocean indicators



---
# Workplan: Objective 3

*Identify opportunities to continue to expand cooperative, constituent-engaged data collection and research to improve the available ecosystem information in the South Atlantic region (e.g., study fleets, cooperative research, citizen science).*

Review existing SAFMC Citizen Science efforts

prioritize potential cooperative data collections based on the analysis of SAFMC management processes most likely to benefit from ecosystem information noted under the previous objective, and outline opportunities for expansion to collect this priority data


---
# Deliverables and Schedule



---
# Coordination

---
.pull-left[
# Discussion
] 

.pull-right[
# Thoughts? Thank you!
]


.footnote[
Slides available at https://sgaichas.github.io/HSpresentations  
Contact: <sgaichas@hydrascientificllc.com>
]

???

