---
title: "Ecosystem Information Review and Strategy Development for More Resilient Fisheries"
subtitle: "SAFMC Project Oversight Team <br />11 September 2025<br /><br /><br /><br /><br /><br /><br /><br /><br /><br /><br />"
author: "Sarah Gaichas, Hydra Scientific LLC"
output:
  xaringan::moon_reader:
    css: ["default", "libs/HS_theme3_508_16by9.css"]
    lib_dir: libs
    nature:
      titleSlideClass: ["right", "bottom", "my-title"]
      ratio: "16:9"
      highlightStyle: githubp
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "libs/macros.js"
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_plugins: ["notes", "search"]
---
class: top, left

```{r setup, include=FALSE}
#<span style="font-weight:normal;"><font size="-0.5">$^1$NOAA NMFS Northeast Fisheries Science Center, Woods Hole, MA, USA; $^2$Ocean Associates Inc, Arlington, VA, USA;</font></span>

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = F,
                      fig.retina = 3,
                      warning = F,
                      message = F,
                      error = F)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
#Plotting and data libraries
library(tidyverse)
theme_set(theme_bw())
library(tidyr)
library(here)
library(sf)
library(patchwork)
library(ggpubr)
library(data.table)
library(forcats)
```

```{r, load_refs, include=FALSE, cache=FALSE}
#trick to downlaod bib file from other repo to keep just 1 up to date

# url <- "https://raw.githubusercontent.com/sgaichas/bluefishdiet/main/FishDiet_EcoIndicators.bib"
#   download.file(url, destfile = "./FishDiet_Ecoindicators.bib")

library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "authoryear",
           longnamesfirst = FALSE,
           max.names = 1,
           style = "markdown")
myBib <- ReadBib("./operationalSOE.bib", check = FALSE)

#,
#           hyperlink = FALSE,
#           dashed = FALSE

```

# Project Overview

1) Review the different types of ecosystem data and information including data products and
decision-support tools (e.g. Ecosystem Status Reports, Climate Vulnerability Assessments, risk
assessments) currently available to the regional management councils and evaluate how they are
used in council management processes;

2) Identify opportunities and methods for incorporating identified ecosystem information into
SAFMC management processes, including inter-jurisdictional management decision-making
processes, and identify practical requirements for successful implementation (e.g. data quality,
frequency of information updates, regional council process consistency, implementation
timelines); and

3) Identify opportunities to continue to expand cooperative, constituent-engaged data collection
and research to improve the available ecosystem information in the South Atlantic region (e.g.,
study fleets, cooperative research, citizen science).

---
# Workplan: Objective 1

*Review the different types of ecosystem data and information including data products and decision-support tools (e.g. Ecosystem Status Reports, Climate Vulnerability Assessments, risk assessments) currently available to the regional management councils and evaluate how they are used in council management processes*

Literature search

Structured interviews

---
# Workplan: Objective 2

*Identify opportunities and methods for incorporating identified ecosystem information into SAFMC management processes, including inter-jurisdictional management decision- making processes, and identify practical requirements for successful implementation (e.g. data quality, frequency of information updates, regional council process consistency, implementation timelines)*

Initial focus on catch specification

Broaden to other processes using review of 3-5 years of Council actions



---
# Workplan: Objective 3

---
# Deliverables

---
# Coordination

---
.pull-left[
# Discussion
] 

.pull-right[
# Thoughts? Thank you!
]


.footnote[
Slides available at https://sgaichas.github.io/HSpresentations  
Contact: <sgaichas@hydrascientificllc.com>
]

???

